Demo arXiv template

### 2023-07-04 tidyindex.qmd
as tf article and restructure rpepo

---
title: "Demo arXiv template"
format:
  arxiv-pdf:
    keep-tex: true  
pdf-engine: pdflatex
author:
  - name: H. Sherry Zhang
    affiliations:
      - name: Monash University
        department: Department of Econometrics and Business Statistics
        address: 29 Ancora Imparo Way, Clayton
        city: Melbourne, VIC
        country: Australia
    orcid: 0000-0002-7122-1463
    email: huize.zhang@monash.edu
    url: https://huizezhangsh.netlify.app/
  - name: Collaborators
    affiliations:
      - name: Monash University
        department: Department of Econometrics and Business Statistics
        address: 29 Ancora Imparo Way, Clayton
        city: Melbourne, VIC
        country: Australia
abstract: |
  - indexes, useful, quantify severity, early monitoring, 
  - A huge number of indexes have been proposed by domain experts, however, a large majority of them are not being adopted, reused, and compared in research or in practice. 
  - One of the reasons for this is the plenty of indexes are quite complex and there is no obvious easy-to-use implementation to apply them to user's data. 
  - The paper describes a general pipeline framework to construct indexes from spatio-temporal data, 
  - This allows all the indexes to be constructed through a uniform data pipeline and different indexes to vary on the details of each step in the data pipeline and their orders. 
  - The pipeline proposed aim to smooth the workflow of index construction through breaking down the complicated steps proposed by various indexes into small building blocks shared by most of the indexes.
  - The framework will be demonstrated with drought indexes as examples, but appliable in general to environmental indexes constructed from multivariate spatio-temporal data
keywords:
  - indexes
  - data pipeline
  - software design
bibliography: bibliography.bib  
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
options(width = 90)
library(tidyverse)
library(latex2exp)
library(patchwork)
library(indri)
library(lmomco)
library(lubridate)
library(SPEI)
library(tsibble)
library(GGally)
library(patchwork)
```

@@

```{r eval = FALSE, echo = TRUE}
DATA %>% 
  var_trans(.method = thornthwaite, Tave = TMED, ..., .new_name = "pet") %>%
  dim_red(diff = prcp - pet) %>%
  aggregate(.var = diff, .scale = 12, .new_name = "agg") %>%
  dist_fit(.method = "lmoms", .var = agg, .dist = DIST) %>%
  augment(.var = agg)
```

@@

```{r fig-toy-example, eval = TRUE, echo = FALSE}
#| fig-align: center
#| fig-width: 10
#| fig-height: 3
#| fig-cap: "Standardised Precipitation Evapotranspiration Index (SPEI) calculated with two distribution fits in Step 3 described above: generalised logistic and pearson III distribution, using the `wichita` data from the package `SPEI`."
library(indri)
library(lmomco)
library(ggplot2)
library(tsibble)
library(SPEI)
library(tidyverse)
library(lubridate)
data(wichita)

dt <- wichita %>%
  mutate(ym = make_yearmonth(YEAR, MONTH), id = 1) %>% 
  init(id = id, time = ym, indicators = PRCP:ACSH) %>%
  var_trans(.method = thornthwaite, Tave = TMED, lat = 37.6475, .new_name = "pet") %>%
  dim_red(diff = PRCP - pet) %>%
  aggregate(.var = diff, .scale = 12) %>%
  dist_fit(.dist = list(loglogistic(), pearsonIII()), .method = "lmoms", .var = .agg) %>%
  augment(.var = .agg)

>>>>

---
title: "Demo arXiv template"
format:
  tandf-pdf:
    keep-tex: true  
pdf-engine: pdflatex
author:
  - name: H. Sherry Zhang
    affiliations:
      - name: Monash University
        department: Department of Econometrics and Business Statistics
        address: 29 Ancora Imparo Way, Clayton
        city: Melbourne, VIC
        country: Australia
    orcid: 0000-0002-7122-1463
    email: huize.zhang@monash.edu
    url: https://huizezhangsh.netlify.app/
  - name: Collaborators
    affiliations:
      - name: Monash University
        department: Department of Econometrics and Business Statistics
        address: 29 Ancora Imparo Way, Clayton
        city: Melbourne, VIC
        country: Australia
abstract: |
  - indexes, useful, quantify severity, early monitoring, 
  - A huge number of indexes have been proposed by domain experts, however, a large majority of them are not being adopted, reused, and compared in research or in practice. 
  - One of the reasons for this is the plenty of indexes are quite complex and there is no obvious easy-to-use implementation to apply them to user's data. 
  - The paper describes a general pipeline framework to construct indexes from spatio-temporal data, 
  - This allows all the indexes to be constructed through a uniform data pipeline and different indexes to vary on the details of each step in the data pipeline and their orders. 
  - The pipeline proposed aim to smooth the workflow of index construction through breaking down the complicated steps proposed by various indexes into small building blocks shared by most of the indexes.
  - The framework will be demonstrated with drought indexes as examples, but appliable in general to environmental indexes constructed from multivariate spatio-temporal data
keywords:
  - indexes
  - data pipeline
  - software design
bibliography: bibliography.bib  
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
options(width = 90)
library(tidyverse)
library(latex2exp)
library(patchwork)
library(tidyindex)
library(lmomco)
library(lubridate)
library(SPEI)
library(tsibble)
library(GGally)
library(patchwork)
```

@@

````
DATA %>% 
  var_trans(.method = thornthwaite, Tave = TMED, ..., .new_name = "pet") %>%
  dim_red(diff = prcp - pet) %>%
  aggregate(.var = diff, .scale = 12, .new_name = "agg") %>%
  dist_fit(.method = "lmoms", .var = agg, .dist = DIST) %>%
  augment(.var = agg)
````

@@

```{r}
knitr::include_graphics(here::here("figures/toy-example-spei.png"))
```

```{r fig-toy-example, eval = FALSE, echo = FALSE}
#| fig-align: center
#| fig-width: 10
#| fig-height: 3
#| fig-cap: "Standardised Precipitation Evapotranspiration Index (SPEI) calculated with two distribution fits in Step 3 described above: generalised logistic and pearson III distribution, using the `wichita` data from the package `SPEI`."
library(tidyindex)
library(lmomco)
library(ggplot2)
library(tsibble)
library(SPEI)
library(tidyverse)
library(lubridate)
data(wichita)

dt <- wichita %>%
  mutate(ym = make_yearmonth(YEAR, MONTH), id = 1) %>% 
  init(id = id, time = ym) %>%
  var_trans(.pet_method = "thornthwaite", .vars = TMED, .lat = 37.6475) %>%
  dim_red(diff = PRCP - pet) %>%
  aggregate(.var = diff, .scale = 12) %>%
  dist_fit(.dist = list(loglogistic(), pearsonIII()), .method = "lmoms", .var = .agg) %>%
  augment(.var = .agg)

### 2023-07-04 tidyindex.tex
as tf article and restructure rpepo

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
]{article}

\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
  \setmathfont[]{Latin Modern Math}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

@@

\usepackage{arxiv}
\usepackage{orcidlink}
\usepackage{amsmath}
\usepackage[T1]{fontenc}
\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[many]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Demo arXiv template},
  pdfauthor={H. Sherry Zhang; Collaborators},
  pdfkeywords={indexes, data pipeline, software design},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{Demo arXiv template}
\author{
\textbf{H. Sherry Zhang}~\orcidlink{0000-0002-7122-1463}\\Department of
Econometrics and Business Statistics\\Monash University\\Melbourne,
VIC\\\href{mailto:huize.zhang@monash.edu}{huize.zhang@monash.edu}\\\\\\
\textbf{Collaborators}\\Department of Econometrics and Business
Statistics\\Monash University\\Melbourne, VIC\\}
\date{}
\begin{document}
\maketitle
\begin{abstract}
\begin{itemize}
\tightlist
\item
  indexes, useful, quantify severity, early monitoring,
\item
  A huge number of indexes have been proposed by domain experts,
  however, a large majority of them are not being adopted, reused, and
  compared in research or in practice.
\item
  One of the reasons for this is the plenty of indexes are quite complex
  and there is no obvious easy-to-use implementation to apply them to
  user's data.
\item
  The paper describes a general pipeline framework to construct indexes
  from spatio-temporal data,
\item
  This allows all the indexes to be constructed through a uniform data
  pipeline and different indexes to vary on the details of each step in
  the data pipeline and their orders.
\item
  The pipeline proposed aim to smooth the workflow of index construction
  through breaking down the complicated steps proposed by various
  indexes into small building blocks shared by most of the indexes.
\item
  The framework will be demonstrated with drought indexes as examples,
  but appliable in general to environmental indexes constructed from
  multivariate spatio-temporal data
\end{itemize}
\end{abstract}
{\bfseries \emph Keywords}
\def\sep{\textbullet\ }
indexes \sep data pipeline \sep 
software design

\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, breakable, borderline west={3pt}{0pt}{shadecolor}, enhanced, frame hidden, interior hidden, sharp corners]}{\end{tcolorbox}}\fi

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{DATA }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{var\_trans}\NormalTok{(}\AttributeTok{.method =}\NormalTok{ thornthwaite, }\AttributeTok{Tave =}\NormalTok{ TMED, ..., }\AttributeTok{.new\_name =} \StringTok{"pet"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{dim\_red}\NormalTok{(}\AttributeTok{diff =}\NormalTok{ prcp }\SpecialCharTok{{-}}\NormalTok{ pet) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{aggregate}\NormalTok{(}\AttributeTok{.var =}\NormalTok{ diff, }\AttributeTok{.scale =} \DecValTok{12}\NormalTok{, }\AttributeTok{.new\_name =} \StringTok{"agg"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{dist\_fit}\NormalTok{(}\AttributeTok{.method =} \StringTok{"lmoms"}\NormalTok{, }\AttributeTok{.var =}\NormalTok{ agg, }\AttributeTok{.dist =}\NormalTok{ DIST) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{augment}\NormalTok{(}\AttributeTok{.var =}\NormalTok{ agg)}
\end{Highlighting}
\end{Shaded}

The pipeline construct allows for multiple \texttt{.scales} and
\texttt{.dist} to be evaluated in \texttt{aggregate()} and
\texttt{dist\_fit()} to compare index under different parameterisations.
The result can then be passed into the \texttt{ggplot2} to crease
visualisation. Figure~\ref{fig-toy-example} compares the SPEI calculated
with two distributions (log-logistic and pearson III).

@@

{\centering \includegraphics{paper_files/figure-pdf/fig-toy-example-1.pdf}

\caption{\label{fig-toy-example}Standardised Precipitation
Evapotranspiration Index (SPEI) calculated with two distribution fits in
Step 3 described above: generalised logistic and pearson III
distribution, using the \texttt{wichita} data from the package
\texttt{SPEI}.}

@@

{\centering \includegraphics[width=1\textwidth,height=0.9\textheight]{../figures/pipeline-steps.png}

@@

{\centering \includegraphics{paper_files/figure-pdf/fig-scale-var-trans-compare-1.pdf}

@@

{\centering \includegraphics{paper_files/figure-pdf/fig-normalising-1.pdf}

@@

\begin{verbatim}
# A tibble: 128,586 x 19
   .idx  .period id              ym  prcp  tmax  tmin  tavg  long   lat name    pet   diff
   <chr>   <dbl> <chr>        <mth> <dbl> <dbl> <dbl> <dbl> <dbl> <dbl> <chr> <dbl>  <dbl>
 1 spei        6 ASN00029~ 1990 Jun   170  29.7  16.2  23.0  142. -15.5 KOWA~  67.8  102. 
 2 spei        6 ASN00029~ 1990 Jun   170  29.7  16.2  23.0  142. -15.5 KOWA~  67.8  102. 
 3 spei        6 ASN00029~ 1990 Jun     0  23.0  11.8  17.4  139. -20.7 MOUN~  30.8  -30.8
 4 spei        6 ASN00029~ 1990 Jun     0  23.0  11.8  17.4  139. -20.7 MOUN~  30.8  -30.8
 5 spei        6 ASN00031~ 1990 Jun   794  25.8  18.1  21.9  146. -16.9 CAIR~  66.3  728. 
 6 spei        6 ASN00031~ 1990 Jun   794  25.8  18.1  21.9  146. -16.9 CAIR~  66.3  728. 
 7 spei        6 ASN00031~ 1990 Jun   504  23.0  13.8  18.4  145. -17.1 WALK~  48.4  456. 
 8 spei        6 ASN00031~ 1990 Jun   504  23.0  13.8  18.4  145. -17.1 WALK~  48.4  456. 
 9 spei        6 ASN00032~ 1990 Jun  1970  23.9  16.4  20.2  146. -17.6 SOUT~  54.5 1915. 
10 spei        6 ASN00032~ 1990 Jun  1970  23.9  16.4  20.2  146. -17.6 SOUT~  54.5 1915. 
# ... with 128,576 more rows, and 6 more variables: .scale <dbl>, .agg <dbl>,
#   .method <chr>, .fitted <dbl>, .dist <chr>, .index <dbl>
\end{verbatim}

@@

{\centering \includegraphics{paper_files/figure-pdf/fig-compute-spatial-1.pdf}

@@

{\centering \includegraphics{paper_files/figure-pdf/fig-compute-temporal-1.pdf}

@@

Data: 
# A tibble: 191 x 13
      id country   life_~1 exp_sch avg_sch gni_pc   sch index0 index1 index2 index3 index4
   <dbl> <chr>       <dbl>   <dbl>   <dbl>  <dbl> <dbl>  <dbl>  <dbl>  <dbl>  <dbl>  <dbl>
 1     1 Switzerl~   0.984   0.917   0.924  0.983 0.920  0.962  0.963   1.67  0.978  0.933
 2     2 Norway      0.973   1       0.867  0.978 0.933  0.961  0.961   1.66  0.969  0.942
 3     3 Iceland     0.964   1       0.918  0.955 0.959  0.959  0.959   1.66  0.963  0.959
 4     4 Hong Kon~   1       0.960   0.815  0.973 0.887  0.952  0.953   1.65  0.986  0.907
 5     5 Australia   0.993   1       0.848  0.936 0.924  0.951  0.951   1.65  0.980  0.932
 6     6 Denmark     0.944   1       0.864  0.967 0.932  0.948  0.948   1.64  0.945  0.937
 7     7 Sweden      0.969   1       0.841  0.952 0.920  0.947  0.947   1.64  0.962  0.928
 8     8 Ireland     0.954   1       0.772  1     0.886  0.945  0.947   1.64  0.952  0.904
 9     9 Germany     0.933   0.945   0.939  0.952 0.942  0.942  0.942   1.63  0.936  0.942
10    10 Netherla~   0.949   1       0.839  0.956 0.919  0.941  0.941   1.63  0.947  0.926
# ... with 181 more rows, 1 more variable: index5 <dbl>, and abbreviated variable name
#   1: life_exp
\end{verbatim}

@@

{\centering \includegraphics{paper_files/figure-pdf/fig-hdi-expr-1.pdf}

>>>>

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
]{interact}

\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

@@

\usepackage{orcidlink}
\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[many]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Demo arXiv template},
  pdfauthor={H. Sherry Zhang; Collaborators},
  pdfkeywords={indexes, data pipeline, software design},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{Demo arXiv template}
\author{H. Sherry
Zhang$\textsuperscript{1}$~\orcidlink{0000-0002-7122-1463}, Collaborators$\textsuperscript{1}$}

\thanks{CONTACT: H. Sherry
Zhang. Email: \href{mailto:huize.zhang@monash.edu}{\nolinkurl{huize.zhang@monash.edu}}. }
\begin{document}
\captionsetup{labelsep=space}
\maketitle
\textsuperscript{1} Department of Econometrics and Business
Statistics, Monash University, Melbourne, VIC, Australia
\begin{abstract}
\begin{itemize}
\tightlist
\item
  indexes, useful, quantify severity, early monitoring,
\item
  A huge number of indexes have been proposed by domain experts,
  however, a large majority of them are not being adopted, reused, and
  compared in research or in practice.
\item
  One of the reasons for this is the plenty of indexes are quite complex
  and there is no obvious easy-to-use implementation to apply them to
  user's data.
\item
  The paper describes a general pipeline framework to construct indexes
  from spatio-temporal data,
\item
  This allows all the indexes to be constructed through a uniform data
  pipeline and different indexes to vary on the details of each step in
  the data pipeline and their orders.
\item
  The pipeline proposed aim to smooth the workflow of index construction
  through breaking down the complicated steps proposed by various
  indexes into small building blocks shared by most of the indexes.
\item
  The framework will be demonstrated with drought indexes as examples,
  but appliable in general to environmental indexes constructed from
  multivariate spatio-temporal data
\end{itemize}
\end{abstract}
\begin{keywords}
\def\sep{;\ }
indexes\sep data pipeline\sep 
software design
\end{keywords}
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[borderline west={3pt}{0pt}{shadecolor}, sharp corners, interior hidden, frame hidden, enhanced, breakable, boxrule=0pt]}{\end{tcolorbox}}\fi

@@

\begin{verbatim}
DATA %>% 
  var_trans(.method = thornthwaite, Tave = TMED, ..., .new_name = "pet") %>%
  dim_red(diff = prcp - pet) %>%
  aggregate(.var = diff, .scale = 12, .new_name = "agg") %>%
  dist_fit(.method = "lmoms", .var = agg, .dist = DIST) %>%
  augment(.var = agg)
\end{verbatim}

The pipeline construct allows for multiple \texttt{.scales} and
\texttt{.dist} to be evaluated in \texttt{aggregate()} and
\texttt{dist\_fit()} to compare index under different parameterisations.
The result can then be passed into the \texttt{ggplot2} to crease
visualisation. \textbf{?@fig-toy-example} compares the SPEI calculated
with two distributions (log-logistic and pearson III).

@@

\includegraphics[width=10in,height=\textheight]{figures/toy-example-spei.png}

@@

{\centering \includegraphics[width=1\textwidth,height=0.9\textheight]{figures/pipeline-steps.png}

@@

{\centering \includegraphics{tidyindex_files/figure-pdf/fig-scale-var-trans-compare-1.pdf}

@@

{\centering \includegraphics{tidyindex_files/figure-pdf/fig-normalising-1.pdf}

@@

\begin{verbatim}
[1] "Checking for missing values (`NA`): all the data must be complete. Input type is vector. Assuming the data are monthly time series starting in January, all regular (non-leap) years."
\end{verbatim}

\begin{verbatim}
# A tibble: 128,586 x 19
   .idx  .period id              ym  prcp  tmax  tmin  tavg  long   lat name    pet   diff
   <chr>   <dbl> <chr>        <mth> <dbl> <dbl> <dbl> <dbl> <dbl> <dbl> <chr> <dbl>  <dbl>
 1 spei        6 ASN00029~ 1990 Jun   170  29.7  16.2  23.0  142. -15.5 KOWA~  85.9   84.1
 2 spei        6 ASN00029~ 1990 Jun   170  29.7  16.2  23.0  142. -15.5 KOWA~  85.9   84.1
 3 spei        6 ASN00029~ 1990 Jun     0  23.0  11.8  17.4  139. -20.7 MOUN~  47.6  -47.6
 4 spei        6 ASN00029~ 1990 Jun     0  23.0  11.8  17.4  139. -20.7 MOUN~  47.6  -47.6
 5 spei        6 ASN00031~ 1990 Jun   794  25.8  18.1  21.9  146. -16.9 CAIR~  80.3  714. 
 6 spei        6 ASN00031~ 1990 Jun   794  25.8  18.1  21.9  146. -16.9 CAIR~  80.3  714. 
 7 spei        6 ASN00031~ 1990 Jun   504  23.0  13.8  18.4  145. -17.1 WALK~  48.0  456. 
 8 spei        6 ASN00031~ 1990 Jun   504  23.0  13.8  18.4  145. -17.1 WALK~  48.0  456. 
 9 spei        6 ASN00032~ 1990 Jun  1970  23.9  16.4  20.2  146. -17.6 SOUT~  70.2 1900. 
10 spei        6 ASN00032~ 1990 Jun  1970  23.9  16.4  20.2  146. -17.6 SOUT~  70.2 1900. 
# i 128,576 more rows
# i 6 more variables: .scale <dbl>, .agg <dbl>, .method <chr>, .fitted <dbl>,
#   .dist <chr>, .index <dbl>
\end{verbatim}

@@

{\centering \includegraphics{tidyindex_files/figure-pdf/fig-compute-spatial-1.pdf}

@@

{\centering \includegraphics{tidyindex_files/figure-pdf/fig-compute-temporal-1.pdf}

@@

Data: 
# A tibble: 191 x 13
      id country  life_exp exp_sch avg_sch gni_pc   sch index0 index1 index2 index3 index4
   <dbl> <chr>       <dbl>   <dbl>   <dbl>  <dbl> <dbl>  <dbl>  <dbl>  <dbl>  <dbl>  <dbl>
 1     1 Switzer~    0.984   0.917   0.924  0.983 0.920  0.962  0.963   1.67  0.978  0.933
 2     2 Norway      0.973   1       0.867  0.978 0.933  0.961  0.961   1.66  0.969  0.942
 3     3 Iceland     0.964   1       0.918  0.955 0.959  0.959  0.959   1.66  0.963  0.959
 4     4 Hong Ko~    1       0.960   0.815  0.973 0.887  0.952  0.953   1.65  0.986  0.907
 5     5 Austral~    0.993   1       0.848  0.936 0.924  0.951  0.951   1.65  0.980  0.932
 6     6 Denmark     0.944   1       0.864  0.967 0.932  0.948  0.948   1.64  0.945  0.937
 7     7 Sweden      0.969   1       0.841  0.952 0.920  0.947  0.947   1.64  0.962  0.928
 8     8 Ireland     0.954   1       0.772  1     0.886  0.945  0.947   1.64  0.952  0.904
 9     9 Germany     0.933   0.945   0.939  0.952 0.942  0.942  0.942   1.63  0.936  0.942
10    10 Netherl~    0.949   1       0.839  0.956 0.919  0.941  0.941   1.63  0.947  0.926
# i 181 more rows
# i 1 more variable: index5 <dbl>
\end{verbatim}

@@

{\centering \includegraphics{tidyindex_files/figure-pdf/fig-hdi-expr-1.pdf}
